Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia

### 2024-02-12 outputs/paper/paper.qmd
modify references and discussion part

# Introduction
Barriers to hiring women exert a significant influence on female labor and present a constraint to increasing economic opportunities for women [@Organizational]. Saudi Arabia is a recognizable example of a country with a low rate of female labor force participation. In 2022, the female labor force participation rate, defined as the percentage of the female population aged 15+ that is employed, is only 27.8% in Saudi Arabia. This is notably lower than the global average of 47.3%, while the male labor force participation rate in Saudi Arabia is relatively high at 80% [@worldbankLaborForce].

The prevailing societal norms in Saudi Arabia, characterized by a low rate of women working outside the home and the customary practice of husbands' guardianship with decision-making authority, contribute to the perception that Saudi men are unwilling to allow their wives to work [@MisperceivedSocialNorm]. This perception is often considered the primary reason behind the low rates of female employment, especially female working outside the home (WWOH). The paper ‘Misperceived Social Norms: Women Working Outside the Home in Saudi Arabia’ by Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott delves into this social norm, as well as Saudi men’s perceptions of opinions regarding WWOH.

Drawing from survey responses and experiments results, they find that a large portion of young married men in Saudi Arabia privately support women working outside the home, and they greatly underestimate the proportion of support by other similar men [@MisperceivedSocialNorm]. The study population is young married husbands in Saudi Arabia. We aim to replicate the estimand that focuses on the proportion of husbands signing up job matching for their wives and the subsequent employment of wives a few months after the experiment. Their findings suggest that men who are given the actual proportion of their peers supporting WWOH are more likely to support their wives’ labor supply, and their wives are more likely to apply for job, attend interviews, be employed and enrol in driving lessons.  Correcting misperceived beliefs manifests positive impacts on females, and the lack of communication emerges as a primary factor contributing to these misperceptions.

@@

## Bias and Weaknesses
The authors, Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott, implemented many modifications to the questions and weights in an attempt to minimize bias in the survey. However, it is acknowledged that bias is an inherent challenge when conducting experiments. The main experiment survey has a participant pool of 500 married men from Riyadh. Since Riyadh is the capital and the largest city in Saudi Arabia, the sample size might be relatively small and potentially biased. The reward offered, typically a five-dollar or twenty-dollar gift card, could be another factor contributing to potential biases considering the GDP in Saudi Arabia. The paper focuses on married men aged 18 to 35 with some level of education. It is noted that, for the majority of individuals, the expected age range for completing college or university education is between 20 and 24. Therefore, setting the age range from 18 to 35 might be a weakness in the experiment.

There is a potential correlation between individuals who did not participate in the follow-up and those who do not support Women Working Outside the Home, also introducing the possibility of bias.

## Next steps
Although there is a customary practice of husbands' guardianship with decision-making authority in Saudi Arabia, gaining insights into the opinions of Saudi women regarding employment and working outside the home is essential for future study. The authors mentioned that ‘expanding the scale and observing how information spreads in networks, and how it affects a large set of outcomes is an important topic for future work’ [@MisperceivedSocialNorm]. Considering the potential biases and the experiment result stating correcting misperceived beliefs increase husbands’ support in WWOH, future research could greatly benefit from the extension of scale to enhance generalizability and actual effects on female employment in Saudi society.

Figure 9 in the original paper reveals that males who often engage in discussions about the WWOH exhibit perceived support only four percent smaller than the actual proportion. For those who rarely or very rarely discuss about the WWOH, their perceived support for WWOH is around forty percent smaller than the actual proportion. If the lack of communication facilitates misperceptions, one would expect a correlation between the frequency of communication and perceptions [@MisperceivedSocialNorm]. Consequently, future studies could delve into strategies in preventing misperceived opinions from influencing social norms, and explore decision-making in the presence of collective misperceptions.

>>>>

# Introduction
Barriers to the employment of women exert a significant influence on female labor, presenting a constraint to increasing economic opportunities for women [@sciencedirectOrganizationalEconomic]. Saudi Arabia is a recognizable example of a country with a low rate of female labor force participation. In 2022, the female labor force participation rate, defined as the percentage of the female population aged 15 and above that is employed, is only 27.8% in Saudi Arabia. This is notably lower than the global average of 47.3%, while the male labor force participation rate in Saudi Arabia remains high at 80% [@worldbankLaborForce].

The prevailing societal norms in Saudi Arabia, characterized by a low rate of women working outside the home and the customary practice of husbands' guardianship with decision-making authority, contribute to the perception that Saudi men are unwilling to permit their wives to work [@MisperceivedSocialNorm]. The paper ‘Misperceived Social Norms: Women Working Outside the Home in Saudi Arabia’ by Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott delves into this social norm, exploring Saudi men’s perceptions of opinions regarding WWOH.

Based on survey responses and experiments results, they find that a large portion of young married men in Saudi Arabia privately support women working outside the home. Importantly, these men greatly underestimate the proportion of support among other similar men [@MisperceivedSocialNorm]. The study population is young married husbands aged 18 to 35 in Saudi Arabia. We aim to replicate the estimand that focuses on the effects of providing husbands with the true proportion of support. Their findings suggest that men who are given the actual proportion of their peers supporting WWOH are more likely to support their wives’ labor supply, and their wives are more likely to apply for job, attend interviews, be employed and enrol in driving lessons. Correcting misperceived beliefs yields positive impacts on females employment, and insufficient communication is associated with an increased likelihood of misperceptions [@MisperceivedSocialNorm].

@@

## Bias, Ethics and Weaknesses
Throughout the survey and experiment, all participants granted permissions for the collection of data. After gathering information, the experimenters ensured that all responses remain confidential and anonymous, so the data we used for replication in this paper was also anonymous, with only the last three digits of their phone numbers retained. As a result, there are no ethical concerns during the process and in the paper.

The authors, Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott, implemented many modifications to the questions and weights in an attempt to minimize bias in the survey. However, it is acknowledged that bias is always a challenge when conducting experiments. The main experiment survey has a participant pool of 500 married men from Riyadh [@MisperceivedSocialNorm]. Since Riyadh is the capital and the largest city in Saudi Arabia, the sample size might be relatively small and potentially biased. Considering the economy in Saudi Arabia and Riyadh, the reward offered, typically a five-dollar or twenty-dollar gift card, could be another factor contributing to the weakness in both the original paper and our paper.

The paper focuses on married men aged 18 to 35 with some level of post-secondary education. It is noted that the enrolment rate among students aged 20-24 in secondary program is relatively high in Saudi Arabia [@EducationGPS], so setting the age range from 18 to 35 might be a weakness in the experiment. There is a potential correlation between individuals who did not participate in the follow-up and those who do not support Women Working Outside the Home, also introducing the possibility of bias.

## Next steps
Although there is a customary practice of husbands' guardianship with decision-making authority in Saudi Arabia, gaining insights into the opinions of Saudi women regarding employment and working outside the home is essential for future study. The authors mentioned that ‘expanding the scale and observing how information spreads in networks, and how it affects a large set of outcomes is an important topic for future work’ [@MisperceivedSocialNorm]. Considering the potential biases and the experiment result claiming that correcting misperceived beliefs increase husbands’ support in WWOH, future research could greatly benefit from the extension of scale to enhance generalizability and actual effects on female employment in Saudi society.

Figure 9 in the original paper reveals that males who often engage in discussions about the WWOH have perceived proportion of support only four percent smaller than the actual proportion. For those who rarely or very rarely discuss about the WWOH, their perceived proportion of support for WWOH is around forty percent smaller than the actual proportion. If the lack of communication facilitates misperceptions, one would expect a correlation between the frequency of communication and perceptions [@MisperceivedSocialNorm]. Consequently, future studies could delve into strategies in preventing misperceived opinions from influencing social norms negatively, and explore decision-making in the presence of collective misperceptions.
